---
title: "11ridge_notes"
author: "Zach White"
date: "10/6/2016"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Library MASS and longley data:
```{r longley}
library(MASS)
data("longley")
```

Scatter plot matrix of all variables with response "Employed" via a formula in the plot.

```{r plots}
pairs(Employed ~ ., data=longley)
```

Correlations among variables.  Which pairs of variables have the highest absolut correlation?
```{r corr}
cor(longley)
```

## Linear Models
Fit the linear model in R. In the formula, `.` means include all predictors in the dataframe.
```{r lm}
longley.lm= lm(Employed ~ ., data=longley)
summary(longley.lm)
```

Note that variables that looked like they had strong relationships with the response have standardized $t$ values that are small, and some coefficients are now negative, when the correlation suggested that at least marginally the coefficient would be positive.

```{r}
summary(lm(Employed ~ scale(longley[,-7]), data=longley))
```

Same model, but now we will use the `scale` function to center and standardize each predictor.


## Ridge Regression

Let's fit the Ridge Regression model using the function `lm.ridge` from MASS.

```{r}
plot(lm.ridge(Employed ~ ., data=longley, lambda=seq(0, 0.1, 0.0001)) )
```

The vector `lambda` is a sequence from 0 (OLS) to .1 incrementing by 0.0001.  This is the same as $k$ from the class notes. The plot shows the ridge coefficients under the scaled predictors.     See how the estimated coefficients switch signs!


How do we choose the ridge parameter $k$ (or lambda)?   One option is to use Cross-validation or the related Generalized Cross Validation (easier to compute)

```{r gcv}
select(lm.ridge(Employed ~ ., data=longley, 
                lambda=seq(0, 0.1, 0.0001)))
```


Refit model with best lambda

```{r}
longley.RReg = lm.ridge(Employed ~ ., data=longley, lambda=.0028)
coef(longley.RReg)
```

or extract from the original object

```{r}

fit = lm.ridge(Employed ~ ., data=longley, lambda=seq(0, 0.1, 0.0001))
best.lambda = as.numeric(names(which.min(fit$GCV)))
coef(fit)[which.min(fit$GCV),]
```

